%  LaTeX template for abstract submission for MCM 2019
%
%  If this paper is for a special (invited) session, put the name of session organizer
%  (first name, family name), and the session title.
%  For plenary talks, put ``Plenary Talk'' for the session title, and Zdravko Botev for the organizer.
%
\insession{Feng}{Li}{Monte Carlo Methods for Large Dependent Data}

% First name and name of the speaker.
\speaker{Feng}{Li}%
%  (put no space here)
% Title of the talk, capitalized.
\title{Bayesian high-dimensional covariate-dependent copula modeling with application to
stocks and text sentiments}

% For each author, give the first name, family name, affiliation, and email.
% Ideally, the affiliation and email should fit on a single line.
% No need to put the full snail mailing address.
\author{Feng}{Li}{Central University of Finance and Economics, China}{feng.li@cufe.edu.cn}



% Type your abstract here.
\abstract{

  Tail-dependence modeling with flexible marginal and copula distributions is widely used
  in financial time series. We extend the covariate dependent copula model of Li and Kang
  (2018) to vine copulas and we allow all pair copula parameters including those that
  describe conditional dependence to be functions of linear predictors. Although vine
  decompositions are not unique, fast heuristic algorithms exist for selecting vine
  structures.  The implementation of these algorithms with covariate dependent copula
  parameters is not straightforward. Instead for our empirical studies we exploit the
  structure of the data themselves. We update the copula component together with the
  marginal components jointly. The joint posterior is not tractable and we use a
  stochastic gradient MCMC within Gibbs sampler, i.e. a Gibbs sampler is used for updating
  the joint parameter components, with each conditional parameter block updated by
  stochastic gradient MCMC algorithm. Our model is applied to financial stock returns with
  text sentiments data.

%  If you have refererences, put them here in a format like below.
%  This can be obtained using BiBTeX with the bib style plain.bst.
%  Note that this must be placed inside the abstract.

}  % End of abstract.
